DATA 621---Business Analytics and Data Mining

### 2020-09-10 HW1/HW1_AF.Rmd
updating text

We are given a data set containing 2,276 records, each of which includes the 
number of wins the team had that year and 15 additional statistics about the 
season. The team wins on average about 50% of the games they play (81 games 
out of 162), with the best season having 146 wins and the worst season having 
0 wins. The data is pretty normally distributed, with most seasons having between
49 and 112 wins (blue lines below). This means that there aren't too many extreme 
seasons where the team won significantly more or significantly less games than 
usual. This serves as a good gut-check for our final predictions; if the 
predicted wins are too high or too low, we know something is probably off.

@@

### What's missing? 
A first look at the data shows that only about 8% of the records have 
a full set of information. The good news is that most of the missing values come 
from what we would expect because they don't happen too often; hit-by-pitch 
(92% missing!), caught stealing (34% missing), and double plays (13% missing). 
Since we have so little hit-by-pitch data, we expect that it doesn't contribute 
to overall wins and will eliminate it from a few of the models we propose. The other
two stats have less than half of the data missing, so we'll need to think of a 
clever way to impute these values. The remaining missing information is from a 
combination of stolen bases and strikeouts (by batters and pitchers). **It seems** 
**completely unreasonable** to have zero strike outs in a season, so this is something 
we'll most certainly have to impute.

@@

### Do the individual stats affect winning?
Intuitively, we expect that Caught stealing, Errors, Hits allowed, Homeruns allowed, 
Strikeouts by batters, and Walks allowed would all have a **negative** impact on
the total wins. In other words, as these stats increase, we would expect that the 
team is less likely to win.

However, when we look at the data for each of these statistics with respect to the 
total number of wins, we don't see a clear relationship across the board. In fact, 
only Errors and Hits allowed seem to have a negative impact on the wins. Caught stealing
and strikeouts by batters appear to be random -- whether the stat for a particular season
is high or low doesn't affect the overall number of wins.

Even more interestingly, Homeruns and Walks allowed have the *opposite* effect; as the number
of homeruns and walks that are allowed from the opposing team increase, the number of wins 
appears to increase as well!

@@

We can look at the same information for the stats that we expect to have a **positive**
effect on wins: Base hits, Doubles, Triples, Homeruns, and Walks by batters, Batters
getting hit by pitches, Stolen bases, Double Plays, and Strikeouts by pitchers.

Many of these stats *do* seem to have an effect on the number of wins. Most noteably,
as the number of base hits and walks increase, the team is more likely to win. We see
a slighter, positive relationship between homeruns, doubles, triples, and stolen bases.
This makes sense when we think about it -- these things tend to happen less often in
games than pure base hits and walks. Double plays and batters hit don't appear to have
an effect on the number of wins either. Once again, this intuitively makes sense because
they are less likely to happen in a game.

@@

@@

>>>>

We are given a data set containing 2,276 records, each of which includes the 
number of wins the team had during a particular year and 15 additional statistics 
about the season. The team wins on average about 50% of the games they play (81 games 
out of 162), with the best season having 146 wins and the worst season having 
0 wins. The data is pretty normally distributed and most seasons have between
49 and 112 wins (blue lines below). This means that there aren't too many extreme 
seasons where the team won significantly more or significantly less games than 
usual. This serves as a good gut-check for our final predictions; if the 
predicted wins are too high or too low, we know something is probably off.

@@

### What's missing? 
A first look at the data shows that only about 8% of the records have a full set
of information. The good news is that most of the missing values come from 
statistics that don't happen too often: hit-by-pitch (92% missing!), caught 
stealing (34% missing), and double plays (13% missing). Since we have so little 
hit-by-pitch data, we expect that it doesn't contribute much to overall wins and will 
eliminate it from a few of the models we propose. The other two stats have less 
than half of the data missing, so we'll need to think of a clever way to fill in 
these values. The remaining missing information is from a combination of stolen 
bases and strikeouts (by batters and pitchers). **It seems completely unreasonable** 
to have zero strike outs in a season, so this is something we'll most certainly have 
to impute.

@@

### Do the individual stats affect winning?
Intuitively, we expect that Caught stealing, Errors, Hits allowed, Homeruns allowed, 
Strikeouts by batters, and Walks allowed would all have a **negative** impact on
the total wins. In other words, as these values increase, we expect that the team 
is less likely to win.

However, when we take a closer look at the data, these relationships aren't obvious. 
In fact, only Errors and Hits allowed seem to have a negative impact on wins. Caught 
stealing and Strikeouts by batters appear to be random; this means that whether the 
stat for a particular season is high or low doesn't affect the overall number of wins.

Even more interestingly, Homeruns allowed and Walks allowed have the *opposite* 
effect; as these stats increase, so do the number of wins!

@@

We can look at the same information for the stats that we expect to have a **positive**
effect on wins: Base hits, Doubles, Triples, Homeruns, Walks, Batters getting hit by 
pitches, Stolen bases, Double Plays, and Strikeouts by pitchers.

Many of these stats *do* seem to have an effect on the number of wins. Most notably,
as the number of base hits and walks increase, so do the number of wins. We see
weaker positive relationships for homeruns, doubles, triples, and stolen bases.
This makes sense when we think about it -- these things tend to happen less often in
games than pure base hits and walks, so they don't have as much of an effect on winning. 
Double plays and batters hit don't appear to have an effect on the number of wins either. 
Once again, this intuitively makes sense because they are less likely to happen in a game.

@@

### Are some stats more skewed than others?

@@

### Are stats correlated?
```{r, echo = FALSE, message = FALSE, warning = FALSE}
